---
title: "Retrieving snow and ice albedo from Landsat data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving snow and ice albedo from Landsat data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to use Landsat surface reflectance, topographic data, and glacier outlines to retrieve snow and ice albedo at the Athabasca Glacier in Canada. We provide the input datasets for 16 August 2020 and guide the user through the processing steps implemented in `SatRbedo`. It is assumed that users have basic knowledge of geographic information systems, satellite remote sensing, and accessing Earth Observation data.

## Load the required packages

First, we load the `SatRbedo` and `{terra}` packages. The latter is used for spatial data manipulation, visualization, and analysis.

```{r setup, message = FALSE}
library(SatRbedo)
library(terra)
```

## Input data

Landsat surface reflectance (Landsat Collection 2 Level-2) and topographic (e.g., SRTM) data for the area of interest can be downloaded from the USGS [EarthExplorer](https://earthexplorer.usgs.gov/). Glacier outlines are available on the [GLIMS](https://www.glims.org/) website.

The `SatRbedo` package provides functions to calculate albedo using two (green and near-infrared), four (blue, green, red, and near-infrared), and five (blue, red, near-infrared, shortwave-infrared 1, and shortwave-infrared 2) spectral bands. These functions also require the solar and view angles. To generate these angles, the [Landsat Angles Creation Tools](https://www.usgs.gov/landsat-missions/solar-illumination-and-sensor-viewing-angle-coefficient-files) can be used. 

After downloading the necessary input data, reproject the digital elevation model (DEM) grid and the glacier outline shapefile to Landsat’s coordinate system, then resample them to 30 m spatial resolution. Crop the DEM, spectral surface reflectance, and angle bands to the area of interest. The tasks can be performed using the `preproc()` function and the `{terra}` package.

This workflow can be applied tor other instruments and sensors, by specifying the spectral bands in the functions.
